---
title: "Lab 10 Homework"
author: "John Abram Flores"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above. For any included plots, make sure they are clearly labeled. You are free to use any plot type that you feel best communicates the results of your analysis.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(here)
library(naniar)
``` 

## Desert Ecology
For this assignment, we are going to use a modified data set on [desert ecology](http://esapubs.org/archive/ecol/E090/118/). The data are from: S. K. Morgan Ernest, Thomas J. Valone, and James H. Brown. 2009. Long-term monitoring and experimental manipulation of a Chihuahuan Desert ecosystem near Portal, Arizona, USA. Ecology 90:1708.
```{r}
deserts <- read_csv(here("lab10", "data", "surveys_complete.csv")) %>% clean_names
```

1. Use the function(s) of your choice to get an idea of its structure, including how NA's are treated. Are the data tidy?  
```{r}
skimr::skim(deserts)
miss_var_summary(deserts)
```

*The data appears tidy, as each variable has its own column, each observation their own row, and each value has its own cell.*
```{r}
deserts
```

2. How many genera and species are represented in the data? What are the total number of observations? Which species is most/ least frequently sampled in the study?

*Dipodomys was the most represented genus in the data.*
```{r}
deserts %>%
  count(genus) %>%
  arrange(desc(n))
```

*Dipodomys merriami appears the most in this study, which is commonly known as Merriam's kangaroo rat.*
```{r}
deserts %>%
  count(genus,species) %>%
  arrange(desc(n))
```

*The most and least frequently sampled species listed below:*
```{r}
deserts %>%
  count(genus,species) %>%
  filter(n==max(n)|n==min(n)) %>%
  arrange(desc(n))
```

3. What is the proportion of taxa included in this study? Show a table and plot that reflects this count.

*From the table, we see that the rodent taxon takes up 98.45% of all taxa taken down in this study.*
```{r}
deserts_proportions <- deserts %>%
  count(taxa) %>%
  mutate(percent_of_observations = (n/34786)*100)
deserts_proportions
```

```{r}
ggplot(data=deserts_proportions, aes(x=taxa, y=percent_of_observations)) +geom_col(color="black") + labs(x="Taxa", y="% of Observations", title="Percentage of taxa in study")
```

*Observing the count of taxa thru a log10 y scale.*
```{r}
deserts %>%
  count(taxa) %>%
  ggplot(aes(x=taxa, y=n)) +geom_col(color="black") +scale_y_log10() +labs(y="# of observations (log10)", x="Taxa", title="Proportion of taxa (log10 of observations") 
```

4. For the taxa included in the study, use the fill option to show the proportion of individuals sampled by `plot_type.`

```{r}
deserts %>% 
		ggplot(aes(x = taxa, fill = plot_type))+
		geom_bar(position = position_fill())+ scale_y_continuous(labels = scales::percent)+coord_flip()
```

5. What is the range of weight for each species included in the study? Remove any observations of weight that are NA so they do not show up in the plot.

```{r}
deserts %>%
  filter(!is.na(weight)) %>%
  ggplot(aes(x=species, y=weight))+geom_boxplot() +coord_flip()
```

6. Add another layer to your answer from #4 using `geom_point` to get an idea of how many measurements were taken for each species.

```{r}
deserts %>%
  filter(!is.na(weight)) %>%
  ggplot(aes(x=species, y=weight))+geom_point() +coord_flip()
```

7. [Dipodomys merriami](https://en.wikipedia.org/wiki/Merriam's_kangaroo_rat) is the most frequently sampled animal in the study. How have the number of observations of this species changed over the years included in the study?

```{r}
deserts %>%
  filter(species=="merriami") %>%
  count(year) %>%
  ggplot(aes(x=year, y=n))+geom_col() +labs(y="# of observations", x="Year", title="Record of Dipodomys merriami throughout the years")
```

8. What is the relationship between `weight` and `hindfoot` length? Consider whether or not over plotting is an issue.

*The line of best fit shows us a positive correlation between weight and hindfoot length, but the data is very condensed in certain regions.*
```{r}
deserts %>%
  filter(!is.na(weight)&!is.na(hindfoot_length)) %>%
  ggplot(aes(x=weight, y=hindfoot_length)) +geom_jitter(size=0.4) +geom_smooth(method=lm, se=F) +labs(y="Hindfoot length")
```

9. Which two species have, on average, the highest weight? Once you have identified them, make a new column that is a ratio of `weight` to `hindfoot_length`. Make a plot that shows the range of this new ratio and fill by sex.

*Rodent albigula has the highest average weight.*
```{r}
deserts %>%
  group_by(taxa,species) %>%
  summarize(avg_weight=mean(weight, na.rm=T)) %>%
  arrange(desc(avg_weight))
```

```{r}
deserts %>%
  filter(species=="albigula" & !is.na(weight) & !is.na(hindfoot_length)) %>%
  mutate(weight_to_hindfoot_ratio=weight/hindfoot_length) %>%
  ggplot(aes(x=species, y=weight_to_hindfoot_ratio, fill=sex)) +geom_boxplot() +labs(title="Ranges of Rodent albigula weight:hindfoot length ratios by sex", y="Weight to Hindfoot ratio")
```

10. Make one plot of your choice! Make sure to include at least two of the aesthetics options you have learned.

*I will compare the weights of rodents for the months of January through June and fill by plot_type.*

```{r}
deserts %>%
  filter(taxa=="Rodent") %>%
  ggplot(aes(x=month, y=weight, fill=plot_type)) +geom_col(position="dodge") +xlim(0,6.75) +labs(title="Rodent weights compared by month and plot type")
```


## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences. 