---
title: "Lab 6 Homework"
author: "John Abram Flores"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(skimr)
```

For this assignment we are going to work with a large data set from the [United Nations Food and Agriculture Organization](http://www.fao.org/about/en/) on world fisheries. These data are pretty wild, so we need to do some cleaning. First, load the data.  

Load the data `FAO_1950to2012_111914.csv` as a new object titled `fisheries`.
```{r}
fisheries <- readr::read_csv(file = "data/FAO_1950to2012_111914.csv")
```

1. Do an exploratory analysis of the data (your choice). What are the names of the variables, what are the dimensions, are there any NA's, what are the classes of the variables?  
*The dimensions of the data frame are shown by dim(), the names and classes of variables are shown by summary(), and we check NA's by running anyNA(), which outputs TRUE, informing us that there are NA's present.*
```{r}
dim(fisheries)
summary(fisheries)
anyNA(fisheries)
```

2. Use `janitor` to rename the columns and make them easier to use. As part of this cleaning step, change `country`, `isscaap_group_number`, `asfis_species_number`, and `fao_major_fishing_area` to data class factor. 

*I first clean up the column names using janitor's clean_names(df) command.*
```{r}
fisheries <- janitor::clean_names(fisheries)
```
*Then, I set the columns that we will observe to factors using as.factor().*
```{r}
fisheries$country <- as.factor(fisheries$country)
fisheries$isscaap_group_number <- as.factor(fisheries$isscaap_group_number)
fisheries$asfis_species_number <- as.factor(fisheries$asfis_species_number)
fisheries$fao_major_fishing_area <- as.factor(fisheries$fao_major_fishing_area)
```

*I double check that the factor setting worked using str(fisheries). The variables we set earlier show up as factors!*
```{r}
str(fisheries)
```

We need to deal with the years because they are being treated as characters and start with an X. We also have the problem that the column names that are years actually represent data. We haven't discussed tidy data yet, so here is some help. You should run this ugly chunk to transform the data for the rest of the homework. It will only work if you have used janitor to rename the variables in question 2!
```{r}
fisheries_tidy <- fisheries %>% 
  pivot_longer(-c(country,common_name,isscaap_group_number,isscaap_taxonomic_group,asfis_species_number,asfis_species_name,fao_major_fishing_area,measure),
               names_to = "year",
               values_to = "catch",
               values_drop_na = TRUE) %>% 
  mutate(year= as.numeric(str_replace(year, 'x', ''))) %>% 
  mutate(catch= str_replace(catch, c(' F'), replacement = '')) %>% 
  mutate(catch= str_replace(catch, c('...'), replacement = '')) %>% 
  mutate(catch= str_replace(catch, c('-'), replacement = '')) %>% 
  mutate(catch= str_replace(catch, c('0 0'), replacement = ''))

fisheries_tidy$catch <- as.numeric(fisheries_tidy$catch)
```

3. How many countries are represented in the data? Provide a count and list their names.
*There are a total of 204 countries, represented by the 204 total names listed below.*
```{r}
levels(fisheries_tidy$country)
```

4. Refocus the data only to include country, isscaap_taxonomic_group, asfis_species_name, asfis_species_number, year, catch.
```{r}
fisheries_tidy_f <- select(fisheries_tidy, country, isscaap_taxonomic_group, asfis_species_name, asfis_species_number, year, catch)
```

5. Based on the asfis_species_number, how many distinct fish species were caught as part of these data?
*By piping the fisheries_tidy data frame into a summarise()  with n_distinct, we see that there are 1551 distinct fish species caught in this data.*
```{r}
fisheries_tidy_f %>%
  summarise(total_distinct = n_distinct(asfis_species_number))
```

6. Which country had the largest overall catch in the year 2000?
*By running a summarize() command to sum all catches by country, we see that China had the largest overal catch in the year 200 with 25889 total fish caught.*
```{r}
fisheries_tidy_f %>%
  select(year,catch,country) %>%
  filter(year==2000) %>%
  group_by(country) %>%
  summarize(year2k_overall_catch = sum(catch, na.rm=TRUE)) %>%
  slice_max(year2k_overall_catch, n=10)
```

7. Which country caught the most sardines (_Sardina pilchardus_) between the years 1990-2000?
*Morocco caught the most sardines between 1990 and 2000 with 7470 total sardines caught.*
```{r}
fisheries_tidy %>%
  filter(asfis_species_name == "Sardina pilchardus") %>%
  filter(between(year,1990,2000))%>%
  group_by(country) %>%
  summarize(sardine_sum = sum(catch, na.rm=TRUE)) %>%
  slice_max(sardine_sum, n=10)
```

8. Which five countries caught the most cephalopods between 2008-2012?
*China, the Republic of Korea, Peru, Japan, and Chile caught the most cephalopods between 2008-2012.*

```{r}
fisheries_tidy %>%
  filter(str_detect(isscaap_taxonomic_group, "Squids")) %>%
  filter(between(year,2008,2012)) %>%
  group_by(country) %>%
  summarize(cephalopod_sum_by_group = sum(catch, na.rm=TRUE)) %>%
  slice_max(cephalopod_sum_by_group, n=10)
```

9. Which species had the highest catch total between 2008-2012? (hint: Osteichthyes is not a species)
*The Theragra chalcogramma, commonly known as the Alaskan pollock, had the highest catch total in 2008-2012 with 41075 total caught.*
```{r}
fisheries_tidy %>%
  filter(!asfis_species_name == "Osteichthyes") %>%
  filter(between(year,2008,2012)) %>%
  group_by(asfis_species_name) %>%
  summarize(total_species_catch = sum(catch, na.rm=TRUE)) %>%
  slice_max(total_species_catch, n=20)
```

10. Use the data to do at least one analysis of your choice.
*With the data, I chose to look at the most caught species of shark, ray, or chimaeras (class of Chondrichthyes) between the years 1972 and 2002. The Elasmobranchii, Rajiformes, and Raja spp are not specific species, and the highest actual species is the Squalus acanthias, or the spiny dogfish, with a total catch of 7522 in the year span.*

```{r}
fisheries_tidy %>%
  filter(str_detect(isscaap_taxonomic_group, "Shark")) %>%
  filter(between(year,1972,2002)) %>%
  group_by(asfis_species_name) %>%
  summarize(total_shark_catch = sum(catch, na.rm = TRUE)) %>%
  slice_max(total_shark_catch, n=10)
```

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.   