---
title: "BIS 15L Midterm 2"
output:
  html_document: 
    theme: spacelab
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(paletteer)
library(naniar)
library(janitor)
```

```{r}
options(scipen=999)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your code should be organized, clean, and run free from errors. Be sure to **add your name** to the author header above. You may use any resources to answer these questions (including each other), but you may not post questions to Open Stacks or external help sites. There are 10 total questions.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean! Your plots should use consistent aesthetics throughout.  

This exam is due by **12:00p on Tuesday, February 22**.  

## Gapminder
For this assignment, we are going to use data from  [gapminder](https://www.gapminder.org/). Gapminder includes information about economics, population, social issues, and life expectancy from countries all over the world. We will use three data sets, so please load all three as separate objects.    

1. population_total.csv  
2. income_per_person_gdppercapita_ppp_inflation_adjusted.csv  
3. life_expectancy_years.csv  

1. (3 points) Once you have an idea of the structure of the data, please make each data set tidy (hint: think back to pivots) and store them as new objects. You will need both the original (wide) and tidy (long) data!  

```{r}
pop <- readr::read_csv(here("midterm2","data", "population_total.csv"))
anyNA(pop)
```

```{r}
pop_tidy <- pop %>%
  pivot_longer(-country, names_to="year", values_to="population")
pop_tidy$year <- as.numeric(pop_tidy$year)
pop_tidy
```

```{r}
income <- readr::read_csv(here("midterm2", "data", "income_per_person_gdppercapita_ppp_inflation_adjusted.csv"))
anyNA(income)
```

```{r}
income_tidy <- income %>%
  pivot_longer(-country, names_to="year", values_to="income")
income_tidy$year <- as.numeric(income_tidy$year)
income_tidy
```

```{r}
life <- readr::read_csv(here("midterm2", "data", "life_expectancy_years.csv"))
```

```{r}
life_tidy <- life %>%
  pivot_longer(-country, names_to="year", values_to="life_expectancy")
life_tidy$year <- as.numeric(life_tidy$year)
life_tidy

miss_var_summary(life_tidy) # Note: there are NA's in the life expectancy data table
```

```{r}
life_tidy %>%
  group_by(country) %>%
  miss_var_summary() %>%
  arrange(desc(n_miss)) #Andorra, Dominica, and the Marshall Islands have NA's
```

2. (1 point) How many different countries are represented in the data? Provide the total number and their names. Since each data set includes different numbers of countries, you will need to do this for each one.  

**In the population data set, there are** _195_ **distinct countries.**
```{r}
pop_tidy %>%
  distinct(country)
```

```{r}
pop_tidy %>%
  summarize(distinct_countries=n_distinct(country))
```

**In the income (GDP per capita) data set, there are** _193_ **distinct countries.**
```{r}
income_tidy %>%
  distinct(country)
```

```{r}
income_tidy %>%
  summarize(distinct_countries=n_distinct(country))
```

**In the life expectancy data set, there are** _187_ **distinct countries.**
```{r}
life_tidy %>%
  distinct(country)
```

```{r}
life_tidy %>%
  summarize(distinct_countries=n_distinct(country))
```

## Life Expectancy  

3. (2 points) Let's limit the data to 100 years (1920-2020). For these years, which country has the highest average life expectancy? How about the lowest average life expectancy?  

**The country with the highest average life expectancy between 1920 and 2020 is Andorra with an average** _79.78_ **years. The country with the lowest is the Central Africa Republic with** _41.76_ **years. There are missing data entries for life expectancy with Andorra, resulting in less 100 years of life expectancies acccoutn for. The country with the second highest average life expectancy (through the full 100 years) is Sweden with** _73.52_**
```{r}
life_tidy %>%
  filter(year<=2020 & year>=1920) %>%
  group_by(country) %>%
  summarize(mean_life_expectancy=mean(life_expectancy, na.rm=T)) %>%
  arrange(desc(mean_life_expectancy)) %>%
  filter(mean_life_expectancy==max(mean_life_expectancy) | mean_life_expectancy==min(mean_life_expectancy) | country=="Sweden") # Note: Andorra has missing data in a range of years
```

```{r}
life_tidy %>%
  filter(year<=2020 & year>=1920) %>%
  group_by(country) %>%
  summarize(mean_life_expectancy=mean(life_expectancy, na.rm=T)) %>%
  arrange(desc(mean_life_expectancy)) %>%
  head(n=5)
```

4. (3 points) Although we can see which country has the highest life expectancy for the past 100 years, we don't know which countries have changed the most. What are the top 5 countries that have experienced the biggest improvement in life expectancy between 1920-2020?  

**The top 5 countries that have experienced the biggest improvement in life expectancy are:** _Kazakhstan, Lithuania, Poland, Ukraine and the Kyrgyz Republic (Kyrgyzstan)._
```{r}
life_tidy %>%
  filter(year<=2020 & year>=1920 & life_expectancy!=is.na(.)) %>%
  group_by(country) %>%
  summarize(life_expectancy_change=max(life_expectancy)-min(life_expectancy)) %>%
  arrange(desc(life_expectancy_change)) %>%
  head(n=5, life_expectancy_change)
```

5. (3 points) Make a plot that shows the change over the past 100 years for the country with the biggest improvement in life expectancy. Be sure to add appropriate aesthetics to make the plot clean and clear. Once you have made the plot, do a little internet searching and see if you can discover what historical event may have contributed to this remarkable change.  

__One historical event that may have contributed to this remarkable change would be the intervention done by the Soviet Union around the 1930s-1960s.__
```{r}
life_tidy %>%
  filter(country=="Kazakhstan" & year<=2020 & year>=1920) %>%
  ggplot(aes(x=year, y=life_expectancy,)) +geom_line(color="blue", size=1.2, alpha=0.75) +theme_linedraw() +labs(x="Year", y="Life Expectancy (years)", title="Life Expectancy in Kazakhstan from 1920-2020") +theme(plot.title=element_text(size=rel(1.5), hjus=0)) +xlim(1920,2020)
```

## Population Growth
6. (3 points) Which 5 countries have had the highest population growth over the past 100 years (1920-2020)?  

**The 5 countries with the highest population growth over the past 100 years are:** _India, China, Indonesia, the United States, and Pakistan._
```{r}
pop_tidy %>%
  filter(year<=2020 & year>=1920) %>%
  group_by(country) %>%
  summarize(population_growth = max(population)-min(population)) %>%
  arrange(desc(population_growth)) %>%
  head(n=10)
```

7. (4 points) Produce a plot that shows the 5 countries that have had the highest population growth over the past 100 years (1920-2020). Which countries appear to have had exponential growth?  

**From face value, China and India have slight hints at exponential growth.**
```{r}
pop_tidy %>%
  filter(year<=2020 & year>=1920) %>%
  filter(country=="India" | country=="China" | country=="Indonesia" | country=="United States" | country=="Pakistan") %>%
  ggplot(aes(x=year, y=population, color=country)) +geom_line(size=1.2) +facet_wrap(~country, ncol=2) +labs(title="Population growth amongst the top 5 countries in 1920-2020", y="Population", x="Year")
```

## Income
The units used for income are gross domestic product per person adjusted for differences in purchasing power in international dollars.

8. (4 points) As in the previous questions, which countries have experienced the biggest growth in per person GDP. Show this as a table and then plot the changes for the top 5 countries. With a bit of research, you should be able to explain the dramatic downturns of the wealthiest economies that occurred during the 1980's.  


**From minimum to maximum values in income, the United Arab Emirates, Qatar, Brunei, Kuwait, and Luxembourg had the greatest increases in GDP per capita.**
```{r}
income_tidy %>% #Sorting from max and min income
  filter(year<=2020 & year>=1920) %>%
  group_by(country) %>%
  summarize(income_growth=max(income)-min(income)) %>%
  arrange(desc(income_growth)) %>%
  head(n=20)
```

**From 1920-2020 incomes, Qatar, Luxembourg, Singapore, Brunei, and Ireland had the greatest increase in GDP per capita.**
```{r}
income_tidy %>%
  pivot_wider(names_from = year, values_from=income, names_prefix = "year_") %>%
  select(country, year_1920, year_2020) %>%
  mutate(income_growth = year_2020-year_1920) %>%
  arrange(desc(income_growth))
```

__We see Qatar reach the top 5 in both lists. On another note, the Iranian Revolution and oil crisis in the 1980s attributed to the dramatic downturns of wealthy economies.__

*Table sorted by max-min income*
```{r}
income_tidy %>%
  filter(country=="United Arab Emirates" | country=="Qatar" | country=="Brunei" | country=="Kuwait" | country == "Luxembourg") %>%
  ggplot(aes(x=year, y=income, color=country)) +geom_line(size=1.2) +labs(x="Year", y="GDP per capita", title="GDP per capita across the top 5 growers (1920-2020), max-min") +theme(plot.title=element_text(size=rel(1.3))) +xlim(1920,2020)
``` 

*Table sorted by subtracting income from 2020 by income from 1920:*
```{r}
income_tidy %>%
  filter(country=="Singapore" | country=="Qatar" | country=="Brunei" | country=="Ireland" | country == "Luxembourg") %>%
  ggplot(aes(x=year, y=income, color=country)) +geom_line(size=1.2) +labs(x="Year", y="GDP per capita", title="GDP per capita across the top 5 growers (1920-2020), 2020-1920") +theme(plot.title=element_text(size=rel(1.3))) +xlim(1920,2020)
```

9. (3 points) Create three new objects that restrict each data set (life expectancy, population, income) to the years 1920-2020. Hint: I suggest doing this with the long form of your data. Once this is done, merge all three data sets using the code I provide below. You may need to adjust the code depending on how you have named your objects. I called mine `life_expectancy_100`, `population_100`, and `income_100`. For some of you, learning these `joins` will be important for your project.  

```{r}
life_expectancy_100 <- life_tidy %>%
  filter(year<=2020 & year>=1920) 
life_expectancy_100
```

```{r}
population_100 <- pop_tidy %>%
  filter(year<=2020 & year>=1920) 
population_100
```

```{r}
income_100 <- income_tidy %>%
  filter(year<=2020 & year>=1920)
income_100
```

```{r}
gapminder_join <- inner_join(life_expectancy_100, population_100, by= c("country", "year"))
gapminder_join <- inner_join(gapminder_join, income_100, by= c("country", "year"))
gapminder_join
```

10. (4 points) Use the joined data to perform an analysis of your choice. The analysis should include a comparison between two or more of the variables `life_expectancy`, `population`, or `income.`  

**Analyzing GDP per capita for countries with lowest average life expectancy**
```{r}
gapminder_join %>%
  group_by(country) %>%
  summarize(mean_life = mean(life_expectancy, na.rm=T)) %>%
  arrange(mean_life) %>%
  head(n=5)
```
```{r}
gapminder_join %>%
  filter(country==c("Central African Republic", "Mali", "Ethiopia", "Sierra Leone", "Guinea-Bissau")) %>%
  ggplot(aes(x=year, y=income, color=country)) +geom_line(size=1.2) +labs(x="Year", y="GDP Per Capita", title="GDP Per Capita in countries with lowest average life expectancies (1920-2020)") +theme(plot.title=element_text(size=rel(1.2)))
```


