---
title: "Lab 14 HW"
output: 
  html_document: 
    keep_md: yes
author: "John Abram Flores"
date: '2022-02-28'
---


### Load the Libraries
```{r}
library(tidyverse)
library(janitor)
library("palmerpenguins")

#install.packages("ggVennDiagram")
library(ggVennDiagram)
library(RColorBrewer)

#install.packages("ggworldcloud")
library(ggwordcloud)


options(scipen=999) #cancels the use of scientific notation for the session
```


### Data
```{r}
superhero_info <- readr::read_csv("data/heroes_information.csv", na = c("", "-99", "-"))
superhero_powers <- readr::read_csv("data/super_hero_powers.csv", na = c("", "-99", "-"))

beachbugs_long <- readr::read_csv("data/beachbugs_long.csv")
```

### 1. 
Clean up the column names (no capitals, not spaces) of `superhero_info`, then use 2 functions to remind yourself of structure of the `superhero_info` data set.

```{r}
superhero_info <- clean_names(superhero_info)
```

```{r}
glimpse(superhero_info)
skimr::skim(superhero_info)
```

### 2.
Are bad guys bigger? Make box-plots of weight by `alignment` (alignment on the x-axis).
```{r}
superhero_info %>%
  filter(!is.na(alignment)) %>%
  ggplot(aes(x=alignment, y=weight, fill=alignment)) +geom_boxplot(alpha=0.5)
```

### 3. 
Now, make a violin plot of weight by `alignment` (alignment on the x-axis). Add some color!
  What information can you observe in the violin plot that was not visible in the boxplot?
```{r}
superhero_info %>%
  filter(!is.na(alignment)) %>%
  ggplot(aes(x=alignment, y=weight, fill=alignment)) +geom_violin(alpha=0.5)
```


### 4. 
Use `alpha = .5` in `geom_boxplot()` and `geom_violin()` to layer both plots on top of one another. What does this tell you about the distribution of weight in "`bad`" guys?
```{r}
superhero_info %>%
  filter(!is.na(alignment)) %>%
  ggplot(aes(x=alignment, y=weight, fill=alignment)) +geom_boxplot(alpha=0.5) +geom_violin(alpha=0.5) +labs(title="Distribution of superhero/villian weight by alignment")
```

### 5. 
Box plots are great for showing how the distribution of a numeric variable (e.g. weight) varies among a categorical variable (e.g. alignment).
  Make your own violin plot with the superhero data. 
  What is your numeric data? 
  **height**
  What is your categorical variable?
  **publisher**


```{r}
superhero_info %>%
  filter(!is.na(height) & !is.na(hair_color)) %>%
  ggplot(aes(x=hair_color, y=height, fill=hair_color)) +geom_boxplot(alpha=0.5) +geom_violin(alpha=0.5) +labs(title="Distribution of superhero/villian height by hair color") +coord_flip()
```

### 6. 
Remind yourself what `beachbugs` looks like. Then generate a heatmap of buglevels by site and year. 
color it with `scale_fill_gradient(low="yellow", high="red")` or colors of your choice. you may find it looks best with one color!
(dont forget, `coord_flip()` is a quick way to improve the look of your plot if you dont like the default orientation)

```{r}
glimpse(beachbugs_long)
```

```{r}
beachbugs_long %>%
  ggplot(aes(x=year, y=site, fill=buglevels)) +geom_tile() +scale_fill_gradient(low="green", high="red")
```

### 7.  
Use the provided code to normalize the beachbug data set. 
Then make a heatmap with the `beachbugs_normalized` data, and use the same color choices as above. Which heatmap do you think is more informative? why?

```{r}
#makes a new column of the highest buglevel observed at each site
beachbugs_w_max <- beachbugs_long %>% 
  group_by(site) %>% 
  mutate(max_buglevel = max(buglevels, na.rm=T)) %>% 
  arrange(site)
beachbugs_w_max

#makes a new table where the buglevels are normalized to the max_buglevel
beachbugs_normalized <- beachbugs_w_max %>% 
  group_by(site) %>% 
  mutate(norm_buglevel = buglevels/max_buglevel) %>% 
  arrange(site,year) %>%
  select(site, year, norm_buglevel)   # you dont have to select(), but I think its a clearer looking table
beachbugs_normalized
```
 
```{r}
beachbugs_normalized %>%
  ggplot(aes(x=year, y=site, fill=norm_buglevel)) +geom_tile() +scale_fill_gradient(low="green", high="red")
```

### 8.
Let's make a venn diagram of `superhero_info`, from 4 questions:
Is their alignment evil? 
Are their eyes red?
Are they male?
Are they bald?

Start by making the 4 vectors, then the list of vectors. The vector for alignment is provided:
### super heros venn
```{r}
# evil
evil_vec <- superhero_info %>%
  filter(Alignment=="bad") %>%
  pull(name)

# red eyes
redeye_vec <- superhero_info %>%
  filter(`Eye color`=="red") %>%
  pull(name)

# male

male_vec <- superhero_info %>%
  filter(Gender=="Male") %>%
  pull(name)

# bald

bald_vec <- superhero_info %>%
  filter(`Hair color`=="No Hair") %>%
  pull(name)

```

Your list of vectors will look something like this:
```{r}
questions_list <- list(evil_vec, redeye_vec, male_vec, bald_vec)
```

### 9. 
Let's make the venn diagram! use the code from lab as a reference. 
```{r}
ggVennDiagram(questions_list, category.names = c("Evil", "Red Eyes", "Male", "Bald"))
```


### 10. Choose one intersection of the venn diagram that is interesting to you. Use dplyr to find the names of the superheros in that intersection. 

```{r}
superhero_info %>%
  filter(Gender!="Male" & Alignment=="bad" & `Eye color`=="red" & `Hair color`!="No Hair") 
```



### 11. Make another venn diagram with the `superhero_info` data. What are your questions? ( At least 2!) 
```{r}
# Tall?
very_tall_vec <- superhero_info %>%
  filter(Height>=200) %>%
  pull(name)
# Has hair?
has_hair_vec <- superhero_info %>%
  filter(`Hair color`!="No Hair") %>%
  pull(name)
#Good?
good_vec <- superhero_info %>%
  filter(Alignment=="good") %>%
  pull(name)

Bruh_vec <- list(very_tall_vec, has_hair_vec, good_vec)
```

```{r}
ggVennDiagram(Bruh_vec, category.names=c("Very tall", "Has Hair", "Good"))
```



### 12.
What are some very common super powers? Lets make a word cloud with the `superhero_powers` data.

Use the provided code to make the frequency table, then make a word cloud with it. 
Remember, you can change `scale_size_area(max_size = 20)` to a different number if the words wont fit!

```{r}
head(superhero_powers)
```

```{r}
power_frequency <- superhero_powers %>% 
  select(-hero_names) %>%
  summarise_all(~(sum(.))) %>%
  pivot_longer(everything(), names_to = "power", values_to = "freq")
power_frequency
```


```{r}
power_frequency %>%
  ggplot(aes(
  label = power, 
  size = freq,
  color = power
  )) +geom_text_wordcloud() +theme_minimal()
```
### 13.  
Who are some very powerful supers? 
 Lets make a different word cloud with the `superhero_powers` data. 
Use the provided code to make the frequency table, then make a word cloud with it.
You can use `hero_names` for the labels, and `sum_powers` for size. 

```{r}

power_quantity <- superhero_powers %>% 
  pivot_longer(-hero_names, names_to = "power", values_to = "yes_or_no") %>% 
  group_by(hero_names) %>% 
  mutate(sum_powers = sum(yes_or_no, na.rm=T)) %>%
  arrange(desc(sum_powers), hero_names, desc(yes_or_no))
power_quantity

#we have to trim down to only the top 50, or it will crowd the image!
power_quantity <- power_quantity %>% 
  ungroup %>%
  distinct(hero_names, sum_powers) %>%
  slice_max(sum_powers, n = 50)
power_quantity

```



```{r}
power_quantity %>%
  ggplot(aes(
  label = hero_names, 
  size = sum_powers,
  color = hero_names
  )) +geom_text_wordcloud() +theme_minimal()
```

## That's it! 🎉
Thanks for coding with us all winter! 
Now go finish your group project :) 

